---
title: "Fastball Pitch Speed and Max Temperature in September"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---


```{r satup, include=FALSE}

library(flexdashboard)
library(tidyverse)
library(lubridate)
library(pitchRx)
library(knitr)
library(kableExtra)
library(weathermetrics)
library(plotly)
library(patchwork)
```

```{r data import}
texas = read_csv(file = "./data/texas.csv")
boston = read_csv(file = "./data/boston.csv")
```

Pitchers for the Texas Rangers (Fig.5a, top) seem to have more consistent four-seam fastball speed as temperature increases compared to visiting pitchers (Fig.5a, bottom), who might be less accustomed to pitching in extreme heat. Boston can be colder than most stadiums in September, which might be associated with decreased pitch speed for vistiting pitchers at cooler temps (Fig.5b, bottom) compared to Red Sox pitchers (Fig.5b, top). 

Column {data-width=500}
-----------------------------------------------------------------------

### Fig. 5a Comparing Four-Seam Fastball Speed of Visting ("bottom") and Rangers ("top") Pitchers

```{r fig.5, message=FALSE, warning=FALSE}
# Texas Rangers Stadium in September
tr9 = ggplot(texas, aes(x = tmax, y = start_speed)) +
  geom_point(aes(color = pitcher_name)) +
  geom_smooth(method = 'lm', color = "black") +
  geom_smooth() +
  facet_wrap(~inning_side) +
  labs(
    title = "@ Texas Rangers",
    x = "Max Temperature (Celcius)",
    y = "Pitch Starting Speed (MPH)",
    caption = "Away Team Pitches in Bottom of Inning"
  ) +
  theme(legend.position = "none")

#Join Graphs

ggplotly(tr9)

```


Column {data-width=500}
-----------------------------------------------------------------------

### Fig. 5b Comparing Four-Seam Fastball Speed of Visting ("bottom") and Red Sox ("top") Pitchers

```{r}
# Boston Red Sox Stadium in September

brs9 = ggplot(boston, aes(x = tmax, y = start_speed)) +
  geom_point(aes(color = pitcher_name)) +
  geom_smooth(method = 'lm', color = "black") +
  geom_smooth() +
  facet_wrap(~inning_side) +
  labs(
    title = "@ Boston Redsox",
    x = "Max Temperature (Celcius)",
    y = "Pitch Starting Speed (MPH)",
    caption = "Pitchers = Colors, Month = September"
  ) +
  theme(legend.position = "none")

ggplotly(brs9)
```

