---
title: "linear_models"
author: "Stephen Lewandowski"
date: "November 29, 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(knitr)
library(kableExtra)
```



```{r import_reduce_data}
pitch_tidy_db <- read.csv("./data/pitch_tidy_db.csv") %>%
  select(des, id, type, start_speed, end_speed, pitch_type, inning_side, inning, count, away, home, pitcher, event, batter_name, pitcher_name, team_name, tfs_zulu, tmax, tmin, tmean, tdmean, heat_index, rh) %>% 
  mutate(home = factor(home),
         away = factor(away),
         pitcher_name = factor(pitcher_name),
         pitch_type = factor(pitch_type),
         pitcher = factor(pitcher),
         type = factor(type),
         event = factor(event),
         team_name = factor(team_name))

# missing `date` from latest data_pipeline
```

## Data tables

### Pitch type frequency
```{r pitch_type table}

pitch_tidy_db %>%
  filter(pitch_type != "NA") %>% 
  group_by(pitch_type) %>% 
  summarise(n = n(),
            mean_s_s = mean(start_speed, na.rm = TRUE),
            sd_s_s = sd(start_speed, na.rm = TRUE),
            mean_e_s = mean(end_speed, na.rm = TRUE),
            sd_e_s = sd(end_speed, na.rm = TRUE),
            mean_inning = mean(inning)) %>%
  top_n(n = 10, wt = n) %>% 
  arrange(desc(n)) %>% 
  kable() %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>% 
   add_header_above(c(" " = 2, "Start Speed" = 2, "End Speed" = 2, " " = 1))
```
Pitch type abbreviations and descriptions are listed at https://www.fangraphs.com/library/pitch-type-abbreviations-classifications/. The top 5 by frequency are:

FF = four-seam fastball  
SL = slider   
FT = two-seam fastball  
CH = changeup  
CU = curveball  
  
I will restrict analysis to four-seam fastballs, of which over 750,000 were thrown in outdoor ballparks during our selected time period.

### Fastballs by pitcher

Most fastballs thrown

```{r number_FF_pitcher}

pitch_tidy_db %>% 
  filter(pitch_type == "FF",
         pitcher_name != "NA") %>% 
  group_by(pitcher_name) %>% 
  summarise(n_ff = n(),
            mean_speed = mean(start_speed, na.rm = TRUE),
            sd_speed = sd(start_speed, na.rm = TRUE)) %>% 
  arrange(desc(n_ff)) %>% 
  slice(1:10) %>% 
  kable() %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) 
  
```



Fastest average speed
```{r fastest_pitchers}

pitch_tidy_db %>% 
  filter(pitch_type == "FF",
         pitcher_name != "NA") %>% 
  group_by(pitcher_name) %>% 
  summarise(mean_speed = mean(start_speed, na.rm = TRUE),
            sd_speed = sd(start_speed, na.rm = TRUE),
            n_ff = n()) %>% 
  arrange(desc(mean_speed)) %>% 
  slice(1:10) %>% 
  kable() %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) 

```

Top speed pitches in our database

```{r}
pitch_tidy_db %>% 
  filter(pitcher_name != "NA") %>% 
  arrange(desc(start_speed, end_speed)) %>% 
  slice(1:10) %>%
  select(start_speed, end_speed, pitch_type, pitcher_name, batter_name, inning, des, home, away, tmax, heat_index) %>% 
  kable() %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) 
```



### Model fitting

```{r}

pitch_tidy_ff <-
  pitch_tidy_db %>% 
  filter(pitch_type == "FF")

```


```{r}
pitch_tidy_ff %>% 
  lm(start_speed ~ heat_index, data = .) 
  broom::tidy() %>% 
  knitr::kable(digits = 3)
```

A one degree increase in heat index increased the start speed of four-seam fastballs by 0.036 mph (p < 0.001).

```{r}
pitch_tidy_ff %>% 
  lm(start_speed ~ tmax, data = .) %>% summary()
  broom::tidy() %>% 
  knitr::kable(digits = 3)
```

A one degree increase in maximum daily temperature increased the start speed of four-seam fastballs by 0.037 mph (p < 0.0001).


### Pitch speed by inning
```{r}
pitch_tidy_ff %>% 
  lm(start_speed ~ inning, data = .) %>%  
  broom::tidy() %>% 
  knitr::kable(digits = 3)

```
A one inning increase is associated with a 0.197 mph increase in the start speed of four-seam fastballs (p < 0.001). This may be due to closers coming into the game and throwing more heat.



### Controlled for pitcher (top 200 fastball throwers)
```{r}

pitch_tidy_ff %>% 
  lm(start_speed ~ inning, data = .) %>%  
  broom::tidy() %>% 
  knitr::kable(digits = 3)
```

