---
title: "Linear Models"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Statistical Analysis: Linear Models

```{r filter ff, include = FALSE}
pitch_tidy_ff = pitch_tidy_db %>% 
  filter(pitch_type == "FF") 
```

We used simple linear regression to examine the effects of daily heat index and daily maximum temperature on the start speed of four-seam fastballs while also considering the effects of the individual pitcher, inning, and ballpark location (home team)(AD1).

### Pitch speed and heat index  
```{r lm_heat}
pitch_tidy_ff %>% 
  lm(start_speed ~ heat_index, data = .) %>% 
  broom::tidy() %>% 
  knitr::kable(digits = 3)
```

A one degree increase in heat index increased the start speed of four-seam fastballs by 0.036 mph (p < 0.001). This is a statistically significant difference, but it is not meaningful.  

### Pitch speed and maximum daily temperature  
```{r}
pitch_tidy_ff %>% 
  lm(start_speed ~ tmax, data = .) %>% summary()
  broom::tidy() %>% 
  knitr::kable(digits = 3)
```

A one degree increase in maximum daily temperature increased the start speed of four-seam fastballs by 0.037 mph (p < 0.0001). This result is nearly identical to the heat index model.  

### Pitch speed by inning  
```{r lm_inning}

pitch_tidy_ff %>% 
  lm(start_speed ~ inning, data = .) %>%  
  broom::tidy() %>% 
  knitr::kable(digits = 3)

```

A one inning increase is associated with a 0.197 mph increase in the start speed of four-seam fastballs (p < 0.001). This may be due to closers coming into the game and throwing more heat. Two-tenths of a mph most likely does not make a difference on a per at-bat basis. 

### Pitch speed by inning, controlled for pitcher (top 200 fastball throwers)  
```{r lm_inning_pitcher}

knitr::opts_chunk$set(
  fig.width = 6,
  fig.height = 6,
  out.width = "90%"
)

pitch_tidy_ff %>% 
  filter(pitcher_name != "NA") %>% 
  group_by(pitcher_name) %>% 
  summarise(n_ff = n()) %>% 
  arrange(desc(n_ff)) %>% 
  slice(1:200) %>%
  ungroup() %>% 
  left_join(., pitch_tidy_ff, by = "pitcher_name") %>% 
  lm(start_speed ~ inning + pitcher_name, data = .) %>%  
  broom::tidy() %>% 
  slice(1:5) %>% 
  knitr::kable(digits = 3) 
  
```

After adjusting for pitcher, the effect of inning is only an increase of 0.002 mph per inning (p = 0.03). The effect of inning will not be carried forward to further weather analysis.  

### Pitch speed and heat index, controlled for pitcher (top 200 fastball throwers)  
```{r fastball throwers}

pitch_tidy_ff %>% 
  filter(pitcher_name != "NA") %>% 
  group_by(pitcher_name) %>% 
  summarise(n_ff = n()) %>% 
  arrange(desc(n_ff)) %>% 
  slice(1:200) %>%
  ungroup() %>% 
  left_join(., pitch_tidy_ff, by = "pitcher_name") %>% 
  lm(start_speed ~ heat_index + pitcher_name, data = .) %>%  
  broom::tidy() %>% 
  slice(1:5) %>% 
  knitr::kable(digits = 3)

```

After adjusting for pitcher, a one degree Celsius increase in heat index is associated with a 0.027 mph increase in fastball speed (p < 0.001). This is less of an effect than the model unadjusted for pitcher.  

### Pitch speed and heat index, controlled for home team 
```{r pitch speed and heat index}

pitch_tidy_ff %>% 
  lm(start_speed ~ heat_index + home, data = .) %>%  
  broom::tidy() %>% 
  knitr::kable(digits = 3)

```

After adjusting for game location, a one degree Celsius increase in heat index is associated with a 0.04 mph increase in fastball speed (p < 0.001). This is still a small effect, but is larger than the other models considered. A 10 degree increase in heat index is associated with a 0.4 mph increase in speed, which could make a difference at the professional level.  
 