---
title: "Report - first draft"
output: 
  github_document:
    toc: true
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_bw() + theme(legend.position = "bottom"))
```

### REPORT- FIRST DRAFT 

#### motivation + related 
```{r}
# Tory 
```

#### Project Questions

Initially, our project aimed to answer the following question:

* What is the association between weather and the speed and type of baseball pitches? We hypothesized that the speed and type of Major League Baseball (MLB) pitches may be negatively impacted by extreme weather.  

To operationalize this, we focused on the following weather data: maximum temperature, relative humidity, and heat index. We then focused in on fastball pitches in particular, hypothesizing that these pitches may be most affected by temperature, humidity, and heat index.  

As our research progressed, and it became clear that overall trends relating max temperature and pitch speed were not apparent, we began investigating if a potential relationship between pitch speed and weather could be modified by if a team *typically* played in an extreme weather condition. In other words, were teams that played the majority of their games in cooler climates, i.e. the Milwaukee Brewers, more affected by heat when they played in hotter climates, for example at the Texas Rangers stadium? This led to a revised question:

* Revised question here.

As part of addressing this question, we proposed the following analytic deliverables (AD): 

* Linear model for the effect of temperature on pitch speed (fastball), adjusted for pitcher and other covariates to be determined. (AD1)
* Graph of variation in proportion of fastball vs off-speed pitches and temperature (AD2)
* Graph of speed of fastballs and temperature (AD3)
* Graph of pitch speed (all pitch types) and humidity (AD4)
* Graph of pitch speed (all pitch types) and heat index (AD5)

#### Data
```{r}
#Stephen + Brennan - intro sentences for data section?
```

## Basball Data

```{r}
#Brennan
```


## Weather data

### PRISM climate model  

We obtained our weather data from the Parameter-elevation Regressions on Independent Slopes Model (PRISM) Daily Spatial Climate Dataset (AN81d) developed by the PRISM Climate Group based at Oregon State University. This is a gridded climate model that assimilates station data from networks around the United States and applies interpolation routines to simulate variations with elevation, coastal effects, temperature inversions, and terrain barriers. The model includes 7 bands of output parameters at a 2.5 arc minute grid resolution (https://developers.google.com/earth-engine/datasets/catalog/OREGONSTATE_PRISM_AN81d).

### Google Earth Engine  

Using the code editor in the Google Earth Engine platform (https://code.earthengine.google.com), we extracted maximum daily temperature (`tmax`), minimum daily temperature (`tmin`), and mean daily dew point temperature (`tdmean`) from April 2015 through October 2018 at each MLB stadium location. We obtained the latitude and longitude coordinates for each ballpark from a Google Fusion Table (https://fusiontables.google.com/DataSource?docid=1EXApOoxEgJUFlbMjUodfxBSWlRvgQNpABeddHqiN#rows:id=1), which we imported to Earth Engine. We used JavaScript code to create arrays that collect each weather parameter mapped over each ballpark location for each day, reduced to the first observation reflecting the grid at the point of interest, which are then flattened to two dimensional tables.    

<img src = "./gee_screen_capture.PNG" height = "500px" width = "400px" />

The image above is a screenshot of the Earth Engine code editor. The map displays maximum daily temperatures across the United States on July 05, 2018 from the PRISM model.

The JavaScript code for downloading PRISM data is provided at (gee.html). A registered Google account is required to access the code editor. 

### Tidy weather data  

Next, we imported the PRISM files for `tmax`, `tmin`, and `tdmean` into R, tidied the data from "wide" to "long", and joined each parameter to a single dataframe by team and date. We created daily heat index values (`heat_index`) from mean daily air temperature and mean daily dew point temperature using the `heat.index` function of the `weathermetrics` package, which applies a heat index equation derived from the US National Weather Service's online heat index calculator. We also used the `dewpoint.to.humidity` function to obtain relative humidity (`rh`).   

### Reproducability  

The raw data files from Earth Engine and the stadium index from Fusion Table are available at the links below:  

"tdmean_ballpark.PRISM.csv": https://drive.google.com/open?id=11Wptk3Eaul1rJEmXB2Xb9T7G_B82i3MK  
"tmin_ballpark.PRISM.csv"  : https://drive.google.com/open?id=1gtjdciCF4zJmrcyjXnUrj-JNDs437MCa  
"tmax_ballpark.PRISM.csv"  : https://drive.google.com/open?id=1AYqLA1Qp3tOOG7hltIN9-aXuYUkF3SSq  
"stadium_index.csv"        : https://drive.google.com/open?id=13mteZpqRSzkeUObmaL6VwYT9s3QNlDOo  

The joined and cleaned weather dataframe is available at:

"weather.csv" : https://drive.google.com/open?id=1jr1_sSAzSn8SVKralTHWGdQWH0BEX08W


## Data Pipeline

Data was joined...

```{r}
#insert data pipeline
```



#### Exploratory Analysis

In order to become familiar with the data, a series of graphs were created to explore potential associations between temperature, relative humidity, heat index, and pitch speed. Initially, our team hypothesized that the proportion of pitches that are fastballs on any given day may decrease during warmer temperatures, and so a graph of the proportion of fasballs and max temperature was created (AD2, below). However, the graph indicated that absolutely no association existed between these variables.

```{r prop. ff and tmax}

# Create proportion of ff to all pitch dataset

propff_db = pitch_tidy_db %>% 
  mutate(ff = str_detect(pitch_type, "FF")) %>% 
  group_by(y, m, d, tmax) %>% 
  summarise(n = n(),
            ff_sum = sum(ff, na.rm = TRUE),
            prop_ff = (ff_sum/n)*100) %>% 
  mutate(prop_ff = round(prop_ff, digits = 1))

# tmax and proportion of ff  

propff_db %>%
  filter(prop_ff != 0) %>% 
  ggplot(aes(x = prop_ff, y = tmax)) +
  geom_point() +
  geom_smooth() +
  labs(
    title = "Proportion of Thrown Pitches that are Fastballs and Max Temperature",
    x = "Proportion of Pitches that are Fastballs",
    y = "Max Temperature (C)"
  )

```

Next, several graphs were created to explore the associations between the speed of fasballs and temperature (AD3). Graphs that depicted fasball speed and temperature overall did not suggest specific associations (not shown), so the data was instead faceted by various variables to explore the potential for variation by items such as pitch outcome, team, and pitch type. As one example below, fastball pitch speed against max temperature was plotted, and then faceted by pitch outcome (such as Strikeout, Walk, Single, etc).

```{r ff speed and temp, by outcome}

pitch_tidy_db %>% 
  filter(pitch_type == "FF") %>% 
  ggplot(aes(x = tmax, y = start_speed)) +
  geom_point() +
  facet_wrap(~event) +
  labs(
    title = "Fastball Pitch Speed and Max Temperature, by Pitch Outcome",
    x = "Max Temperature (C)",
    y = "Pitch Starting Speed (MPH)"
  )

```

Interestingly, we can see from the increase in points on the right side of facets such as "Double," "Flyout," "Lineout," "Home Run," and other outcomes that pitch outcome may be affected by temperature. It may be possible that when temperatures are higher, pitches that are slightly slower are more likely to result in hits.  

In further exploratory analysis, fastball speed and tempearture was also faceted by home team name (shown below), inning, and fowl/ball/strike status, and a graph of max temperature and pitch speed for all pitches, faceted by pitch type, was created (all not shown). There were some variations by inning and home team, and after discussion, we decided this might in fact be representative of changes in pitcher, and differences in whether or not a pitcher from a specific team typically played in warmer or colder weather. We decided to perform further analysis on this hypothesis; plots are shown in the Additional Analysis: Plots section of this report. 

```{r ff speed and temp, by home team name}

pitch_tidy_db %>% 
  filter(pitch_type == "FF") %>% 
  ggplot(aes(x = tmax, y = start_speed)) +
  geom_point() +
  facet_wrap(~team_name) +
  labs(
    title = "Fastball Pitch Speed and Max Temperature, by Home Team",
    x = "Max Temperature (C)",
    y = "Pitch Starting Speed (MPH)"
  )

```


```{r}
#Nicole 
```

#### Additional Anlaysis: Plots

The plot of fastball pitch speed by max temperature, faceted by home team, suggested that temperature may affect pitches thrown at several stadiums. Texas Rangers stadium was selected as an example becuase it had a number of pitches with decreased speed at higher temperatures, and a Shiny plot showing associations between pitch speed and temperature at the Texas Rangers's stadium was created. A user can select the visiting team to see how its pitch speed is affected by temperature when playing @ the Texas Rangers.

```{r}
#Brennan - add shiny plot in here
```

Further analysis restricted pitch type to the four-seam fastball, and facetted by top and bottom of the inning, to see if Rangers' pitchers were less effected by heat than visiting pitchers. Additionally, a team from a colder climate, the Boston Red Sox, was selected for the same analysis. Month was restricted to September for two reasons:  (1) during this month temperatures are in greater flux across the United States, and it may be reasonable to assume Texas is much hotter than other parts of the US and Boston is much colder, and (2) it has been suggested anecdotally that baseball pitchers may be affected most by changes in temperature, rather than by absolute heat or cold, particularly if they are not acclimatized to that temperature.  

The resulting plot is shown below. 

```{r ff speed and tmax in sept, rangers and red sox stadiums}
# Texas Rangers Stadium in September

tr9 = pitch_tidy_db %>% 
  filter(pitch_type == "FF" & team_name == "Texas Rangers" & m == "09") %>% 
  ggplot(aes(x = tmax, y = start_speed)) +
  geom_point(aes(color = pitcher_name)) +
  geom_smooth(method = 'lm', color = "black") +
  geom_smooth() +
  facet_wrap(~inning_side) +
  labs(
    title = "Fastball Pitch Speed and Max Temperature at Texas Rangers Stadium", subtitle = "Pitchers = Colors, Month = September",
    x = "Max Temperature (C)",
    y = "Pitch Starting Speed (MPH)"
  ) +
  theme(legend.position = "none")

# Boston Red Sox Stadium in September

brs9 = pitch_tidy_db %>% 
  filter(pitch_type == "FF" & team_name == "Boston Red Sox" & m == "09") %>% 
  ggplot(aes(x = tmax, y = start_speed)) +
  geom_point(aes(color = pitcher_name)) +
  geom_smooth(method = 'lm', color = "black") +
  geom_smooth() +
  facet_wrap(~inning_side) +
  labs(
    title = "Fastball Pitch Speed and Max Temperature at Boston Red Sox Stadium", subtitle = "Pitchers = Colors, Month = September",
    x = "Max Temperature (C)",
    y = "Pitch Starting Speed (MPH)"
  ) +
  theme(legend.position = "none")

#Join Graphs

library(patchwork)
tr9 + brs9

```

In the plot, we can see that in Arlington, Texas, at the Texas Rangers's Stadium, the Rangers appear to be less effected by variations in temperature, whereas visiting pitchers (the bottom of the inning) have a decrease in fastball speed at high temperatures. However, it should be noted that this trend may be driven by one visiting pitcher with particularly slow pitches in hotter weather. In Boston, it appears that visiting pitchers have decreases in fastball speed at colder temperatures, whereas Red Sox pitchers are more consistent regardless of temperature. 

```{r}
#Nicole 
```

#### Additional Analysis: Models
```{r}
#Stephen
```

#### Discussion
```{r}
#All of us
```

