---
title: "Report - first draft"
output: github_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### REPORT- FIRST DRAFT 

#### motivation + related 
Advanced statistical analytics are central to developing teams, evaluating players, and informing in-game decisions throughout professional sports. The growing influence of sport analytics is arguably most evident in Major League Baseball, where teams that prioritize sabermetrics have won the last three World Series. Baseball is uniquely suited to statistical analysis, as its components – pitching, hitting, and fielding – are discrete events that result in unambiguous decisions, as classified by the official game scorers, and can be attributed to specific players. Baseball also generates a substantial amount of data; the 30 MLB teams play a combined 2,430 games in the regular season, during which over 900,000 pitches are thrown across more than 165,000 at-bats. 

Sabermetric analysts sift through this exhaustive amount of data to identify competitive advantages among increasingly similar, dominant players. The arms race between pitchers and hitters has elevated the game, as pitchers are throwing harder and batters are hitting more home runs than at any point in MLB history. 

In addition to assessing player matchups, analysts aim to identify external factors that could influence players’ performances, especially weather.  Over the course of a six-month MLB season, teams expect to play in a range of weather extremes, from 100-degree days in Arlington, Texas to snowy conditions in Milwaukee, Wisconsin in the early spring. A growing body of sabermetric research has found that game-day weather affects umpiring decisions, pitcher ball control, and distance of hits (and thereby number of homeruns).  


```{r}
#The 
complete_db %>% 
  group_by(y) %>% 
  count()

```

#### questions
```{r}
# All of us - take from project proposal 
```

#### data
```{r}
#Stephen + Brennan 
```

#### exploratory analysis 
```{r}
#Jenni + Nicole 
```

#### additional anlaysis: plots
```{r}
#Jenni + Nicole 
```

#### additional analysis: models
```{r}
#Stephen
```

#### discussion
```{r}
#All of us
```

