---
title: "weather_data"
author: "Stephen Lewandowski"
date: "November 9, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```


Stadium index.
Geocoded source: https://fusiontables.google.com/DataSource?docid=1EXApOoxEgJUFlbMjUodfxBSWlRvgQNpABeddHqiN#rows:id=1


```{r stadium_index}
stadium_index <-
  read.csv("./data/stadium_index.csv")
```

Import max and min daily temperatures and mean daily dew points for all MLB stadium locations from April 2015 to October 2018.

```{r import_PRISM_parameters}

tmax <-
  read.csv("./data/tmax_ballpark_PRISM.csv") %>% 
    janitor::clean_names() %>% 
    select(-c("system_index", "geo")) %>% 
    select(team_name, everything()) %>% 
    gather(key = date, value = tmax, "x20150401":"x20181030")
  
tmin <-
  read.csv("./data/tmin_ballpark_PRISM.csv") %>% 
    janitor::clean_names() %>% 
    select(-c("system_index", "geo")) %>% 
    select(team_name, everything()) %>% 
    gather(key = date, value = tmin, "x20150401":"x20181030")

tdmean <-
  read.csv("./data/tdmean_ballpark_PRISM.csv") %>% 
    janitor::clean_names() %>% 
    select(-c("system_index", "geo")) %>% 
    select(team_name, everything()) %>% 
    gather(key = date, value = tdmean, "x20150401":"x20181030")

```

Join tmax, tmin and tdmean

```{r weather_tidy}

weather <- as.tibble(
    full_join(tmax, tmin, by = c("team_name", "date")) %>% 
    full_join(., tdmean, by = c("team_name", "date")) %>% 
    mutate(date = str_replace(date, "x", ""),
           date = as.Date(date, format = "%Y%m%d"))
  )

weather

```





### Google Earth Engine Javascrip code

Example for PRISM mean daily dew point temperature:

// Load a Fusion Table from the ID using the FeatureCollection constructor.

var ballpark = ee.FeatureCollection("ft:1EXApOoxEgJUFlbMjUodfxBSWlRvgQNpABeddHqiN"); 
 
var usa = ee.FeatureCollection("ft:1ksBc_-cWnek37HFTMr0FXfBtOETBeYYbB1YeSg");
   
// Load PRISM input imagery.
var tmax = ee.ImageCollection(PRISM)
.filterDate('2015-04-01', '2018-10-31')
.select('tmax');


var tmin = ee.ImageCollection(PRISM)
.filterDate('2015-04-01', '2018-10-31')
.select('tmin');


var tdmean = ee.ImageCollection(PRISM)
.filterDate('2015-04-01', '2018-10-31')
.select('tdmean');


/////////////////////////////////////////////////////////

//Create a collection of triplets, {imageID, ballpark, first}:

var triplets = tdmean.map(function(image) {
  return image.select('tdmean').reduceRegions({
    collection: ballpark.select(['Team Name']), 
    reducer: ee.Reducer.first(), 
    scale: 1000
  }).filter(ee.Filter.neq('first', null))
    .map(function(f) { 
      return f.set('imageId', image.id());
    });
}).flatten();

print(triplets.first());  //properties: ballpark, imageId, first

//////////////////////////////////////////////////////////////////////////

// Format a table of triplets into a 2D table of rowId x colId.
var format = function(table, rowId, colId) {
  // Get a FeatureCollection with unique row IDs.
  var rows = table.distinct(rowId);
  // Join the table to the unique IDs to get a collection in which
  // each feature stores a list of all features having a common row ID. 
  var joined = ee.Join.saveAll('matches').apply({
    primary: rows, 
    secondary: table, 
    condition: ee.Filter.equals({
      leftField: rowId, 
      rightField: rowId
    })
  });

  return joined.map(function(row) {
      // Get the list of all features with a unique row ID.
      var values = ee.List(row.get('matches'))
        // Map a function over the list of rows to return a list of
        // column ID and value.
        .map(function(feature) {
          feature = ee.Feature(feature);
          return [feature.get(colId), feature.get('first')];
        });
      // Return the row with its ID property and properties for
      // all matching columns IDs storing the output of the reducer.
      // The Dictionary constructor is using a list of key, value pairs.
      return row.select([rowId]).set(ee.Dictionary(values.flatten()));
    });
};


var table1 = format(triplets, 'Team Name', 'imageId');
var desc2 = 'tdmean';

Export.table.toDrive({
  collection: table1,
  description: desc2, 
  fileNamePrefix: desc2,
  fileFormat: 'CSV'
});