---
title: "comfort_graphs"
author: "Nicole Comfort"
date: "11/19/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_bw() + theme(legend.position = "bottom"))

```

## Nicole's EDA looking at humidity and heat index as variables of interest

# Heat Index

For people working outdoors in hot weather, both air temperature and humidity affect how hot they feel. The "heat index" is a single value that takes both temperature and humidity into account. The higher the heat index, the hotter the weather feels, since sweat does not readily evaporate and cool the skin. The heat index is a better measure than air temperature alone for estimating the risk to workers from environmental heat sources.

```{r heat index EDA}

# Examining missing data
complete_db %>% 
  group_by(team_name) %>% 
  filter(is.na(heat_index)) %>% # we are missing pitch speed for 62,971 pitches 
  count() %>% 
  ungroup()# see if the missing data are (roughly) evenly distributed among the teams, Cubs are missing the most but it still seems pretty even 

# examine distribution of heat index
complete_db %>% 
  ggplot(aes(x = heat_index)) +
  geom_histogram()
# pretty much normally distributed overall

# distribution by stadium
complete_db %>% 
  ggplot(aes(x = heat_index, fill = team_name)) + 
  geom_density(alpha = .4, adjust = .5)
# we can see that there are some teams that have a broader distribution; places like San Diego the heat index is pretty stable; we can break up by regions in other analyses 

# numerical summary
skim(complete_db$heat_index)

heat_index_df =
  complete_db %>%
  group_by(team_name, m) %>%
  summarize(mean_heat_index = mean(heat_index, na.rm = TRUE),
            median_heat_index = median(heat_index),
            sd_heat_index = sd(heat_index))

# plot the above table/info 
heat_index_df %>% 
  ggplot(aes(x = m, y = mean_heat_index, color = team_name)) +
  geom_point() + geom_line()
# why doesn't geom_line work???

# could also plot as a box plot of heat index for each team / US region 

# the teams with the greatest range/variability in heat index are the Minnesota Twins, Washington Nationals, Cincinatti Reds, New York Yankees, and Kansas City Royals

# the teams with the lowest range/variability in heat index (excluding values of 0.00) are the Los Angeles Angels of Anaheim, Detroit Tigers, Chicago White Sox, St. Louis Cardinals

# maybe we can do comparisons within certain teams, for example how does someone on the Washington Nationals perform in high vs. low heat index 

complete_db %>% 
  filter(team_name == "Washington Nationals") %>% 
  ggplot(aes(x = m, y = start_speed)) +
  geom_boxplot()
# tried adding geom_jitter but there are too many points 
# more outliers at the warmer months with hotter heat index, but overall start speed looks the same 
```


```{r start pitch speed & heat index, continuous, by MLB region}

# basic scatterplot of pitch speed (all pitches) vs. heat index
complete_db %>% 
  ggplot(aes(x = heat_index, y = start_speed)) +
  geom_point(alpha = .3)
# not informative

# basic scatterplot of pitch speed vs. heat index, facet by pitch type
complete_db %>% 
  ggplot(aes(x = heat_index, y = start_speed)) +
  geom_point(alpha = .3) +
  facet_wrap(~pitch_type)
# too many points... let's use geom_hex
complete_db %>% 
  ggplot(aes(x = heat_index, y = start_speed, color = team_name)) +
  geom_hex() +
  facet_wrap(~pitch_type)
# very ugly and messy

# separate by MLB region
# create new region variable
complete_db =
  dplyr::mutate(complete_db,
                mlb_region =
                  ifelse(team_name == "Baltimore Orioles" | team_name == "Boston Red Sox" |
                           team_name == "New York Yankees" | team_name == "Tampa Bay Rays", "al_east",
                         
                  ifelse(team_name == "Atlanta Braves" | team_name == "Miami Marlins" |
                           team_name == "New York Mets" | team_name == "Philadelphia Phillies" |
                           team_name == "Washington Nationals", "nl_east",
                         
                  ifelse(team_name == "Chicago White Sox" | team_name == "Cleveland Indians" |
                           team_name == "Detroit Tigers" | team_name == "Kansas City Royals" |
                           team_name == "Minnesota Twins", "al_central",
                         
                  ifelse(team_name == "Chicago Cubs" | team_name == "Cincinnati Reds" |
                           team_name == "Milwaukee Brewers" | team_name == "Pittsburgh Pirates" |
                           team_name == "St. Louis Cardinals", "nl_central",
                         
                  ifelse(team_name == "Houston Astros" | team_name == "Los Angeles Angels of Anaheim" |
                           team_name == "Oakland Athletics" | team_name == "Seattle Mariners" |
                           team_name == "Texas Rangers", "al_west",
                         
                  ifelse(team_name == "Arizona Diamondbacks" | team_name == "Colorado Rockies" |
                           team_name == "Los Angeles Dodgers" | team_name == "San Diego Padres" |
                           team_name == "San Francisco Giants", "nl_west", "")))))))

# heat index distribution by MLB region (now that I made the region variable)
complete_db %>% 
  ggplot(aes(x = heat_index, fill = mlb_region)) + 
  geom_density(alpha = .4, adjust = .5)

complete_db %>% 
  ggplot(aes(x = heat_index)) + 
  geom_density(alpha = .4, adjust = .5) +
  facet_wrap(~mlb_region)


# basic scatterplot of pitch speed vs. heat index, facet by pitch type, color by MLB region
complete_db %>% 
  ggplot(aes(x = heat_index, y = start_speed, color = mlb_region)) +
  geom_point(alpha = .3) +
  facet_wrap(~pitch_type)
# the line looks pretty flat; appears there's no association between heat index and pitch speed.


# look at pitch speed (all pitches) vs. heat index
complete_db %>% 
  ggplot(aes(x = heat_index, y = start_speed)) +
  geom_hex() + 
  geom_smooth()
# again, looks like no association. maybe it dips slightly at the highest heat index, like above 35? Let me try to categorize heat index next

```

```{r pitch speed & heat index, continuous, by US region}

# create new variable for US region
complete_db = 
   dplyr::mutate(complete_db,
                  US_region =
                  ifelse(team_name == "Baltimore Orioles" |
                           team_name == "Boston Red Sox" |
                           team_name == "New York Yankees" |
                           team_name == "New York Mets" |
                           team_name == "Tampa Bay Rays" |
                           team_name == "Philadelphia Phillies" |
                           team_name == "Pittsburgh Pirates" |
                           team_name == "Washington Nationals", "northeast",
                  ifelse(team_name == "Chicago Cubs" |
                           team_name == "Chicago White Sox" |
                           team_name == "Cleveland Indians" |
                           team_name == "Detroit Tigers" |
                           team_name == "Kansas City Royals" |
                           team_name == "Minnesota Twins" |
                           team_name == "Cincinnati Reds" |
                           team_name == "Milwaukee Brewers" |
                           team_name == "St. Louis Cardinals", "midwest",
                  ifelse(team_name == "Texas Rangers" |
                           team_name == "Houston Astros", "southwest",
                  ifelse(team_name == "Los Angeles Angels of Anaheim" |
                           team_name == "Los Angeles Dodgers" |
                           team_name == "Oakland Athletics" |
                           team_name == "Seattle Mariners" |
                           team_name == "San Diego Padres" |
                           team_name == "San Francisco Giants" |
                           team_name == "Colorado Rockies", "west",
                  ifelse(team_name == "Atlanta Braves", "southeast", ""))))))

# heat index distribution by US region 
complete_db %>% 
  ggplot(aes(x = heat_index, fill = US_region)) + 
  geom_density(alpha = .4, adjust = .5)
# southwest followed by southeast has highest heat index, west is most stable/pleasant

complete_db %>% 
  ggplot(aes(x = heat_index)) + 
  geom_density(alpha = .4, adjust = .5) +
  facet_wrap(~US_region)
# west has pretty normal distribution, southeast and southwest skewed 

# same but by US region
complete_db %>% 
  ggplot(aes(x = heat_index, y = start_speed, color = US_region)) +
  geom_point(alpha = .3) +
  facet_wrap(~pitch_type)
# fastball affected by higher heat index??
# looks like fastball speed decreases with higher heat index (at least in the southwest and west)

```


```{r start pitch speed & heat index, categorical}

# put heat index into meaningful(?) categories - might need Stephen's input on this
# try above and below 35
# then try quartiles (based on skim summary statistics) 

library(patchwork)

# look at pitch speed (all pitches) vs. heat index
plot_1 = complete_db %>% 
  filter(heat_index < 35) %>% 
  ggplot(aes(x = heat_index, y = start_speed)) +
  geom_hex() + 
  geom_smooth(color = "blue")

plot_2 = complete_db %>% 
  filter(heat_index >= 35) %>% 
  ggplot(aes(x = heat_index, y = start_speed)) +
  geom_point() + 
  geom_smooth(color = "red")
# run a linear regression to see if it increases at these extreme heat indexes? 

plot_1 / plot_2

```


# Relative Humidity 

Relative humidity is a measure of the amount of moisture in the air. Sweat does not evaporate as quickly when the air is moist as it does in a dry climate. Since evaporation of sweat from the skin is one of the ways the human body cools itself on a hot day, high humidity reduces our natural cooling potential and we feel hotter. Low humidity can also be a problem for outdoor workers in hot, desert-like climates. Sweat evaporates very rapidly in low humidity, which can lead to severe dehydration if a person does not drink enough water throughout the day.

```{r start pitch speed & humidity}

# scatterplot looking at pitch speed vs. relative humidity 
complete_db %>% 
  ggplot(aes(x = rh, y = start_speed, color = mlb_region)) +
  geom_point(alpha = .3) +
  facet_wrap(~pitch_type)

# maybe EP, FO, KN, PO, and SC are increasing speed with increasing relative humidity? Could change the scale on the start_speed to better assess

# also, what pitch type is EP and SC? It looks like the only pitcher(s) throwing these belong to the american league central division 

complete_db %>% 
  ggplot(aes(x = rh, y = start_speed, color = US_region)) +
  geom_point(alpha = .3) +
  facet_wrap(~pitch_type)
# U-shaped for fastball? 

# look at pitch speed (all pitches) vs. relative humidity
complete_db %>% 
  ggplot(aes(x = rh, y = start_speed)) +
  geom_hex() + 
  geom_smooth()
# again, looks like no association 


```

## Examine across years? 

```{r}

# complete_db %>% 
#   group_by(y) %>% 
#   summarize(mean = mean(start_speed))

```

