---
title: "Tables and Models"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(knitr)
library(kableExtra)
```



```{r import_reduce_data}
pitch_tidy_db <- read.csv("./data/pitch_tidy_db.csv") %>%
  select(des, id, type, start_speed, end_speed, pitch_type, inning_side, inning, count, away, home, pitcher, event, batter_name, pitcher_name, team_name, tfs_zulu, tmax, tmin, tmean, tdmean, heat_index, rh) %>% 
  mutate(home = factor(home),
         away = factor(away),
         pitcher_name = factor(pitcher_name),
         pitch_type = factor(pitch_type),
         pitcher = factor(pitcher),
         type = factor(type),
         event = factor(event),
         team_name = factor(team_name))

# missing `date` from latest data_pipeline
```

## Data tables

### Pitch type frequency
```{r pitch_type table}

pitch_tidy_db %>%
  filter(pitch_type != "NA") %>% 
  group_by(pitch_type) %>% 
  summarise(n = n(),
            mean_s_s = mean(start_speed, na.rm = TRUE),
            sd_s_s = sd(start_speed, na.rm = TRUE),
            mean_e_s = mean(end_speed, na.rm = TRUE),
            sd_e_s = sd(end_speed, na.rm = TRUE),
            mean_inning = mean(inning)) %>%
  top_n(n = 10, wt = n) %>% 
  arrange(desc(n)) %>% 
  kable() %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>% 
   add_header_above(c(" " = 2, "Start Speed" = 2, "End Speed" = 2, " " = 1))
```
Pitch type abbreviations and descriptions are listed at https://www.fangraphs.com/library/pitch-type-abbreviations-classifications/. The top 5 by frequency are:

FF = four-seam fastball  
SL = slider   
FT = two-seam fastball  
CH = changeup  
CU = curveball  
  
I will restrict analysis to four-seam fastballs, of which over 750,000 were thrown in outdoor ballparks during our selected time period.

### Fastballs by pitcher

Most fastballs thrown

```{r number_FF_pitcher}

pitch_tidy_db %>% 
  filter(pitch_type == "FF",
         pitcher_name != "NA") %>% 
  group_by(pitcher_name) %>% 
  summarise(n_ff = n(),
            mean_speed = mean(start_speed, na.rm = TRUE),
            sd_speed = sd(start_speed, na.rm = TRUE)) %>% 
  arrange(desc(n_ff)) %>% 
  slice(1:10) %>% 
  kable() %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) 
  
```



Fastest average speed
```{r fastest_pitchers}

pitch_tidy_db %>% 
  filter(pitch_type == "FF",
         pitcher_name != "NA") %>% 
  group_by(pitcher_name) %>% 
  summarise(mean_speed = mean(start_speed, na.rm = TRUE),
            sd_speed = sd(start_speed, na.rm = TRUE),
            n_ff = n()) %>% 
  arrange(desc(mean_speed)) %>% 
  slice(1:10) %>% 
  kable() %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) 

```

Top speed pitches in our database

```{r table_pitch_speed}
pitch_tidy_db %>% 
  filter(pitcher_name != "NA") %>% 
  arrange(desc(start_speed, end_speed)) %>% 
  slice(1:10) %>%
  select(start_speed, end_speed, pitch_type, pitcher_name, batter_name, inning, des, home, away, tmax, heat_index) %>% 
  kable() %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) 
```




## Statistical Analysis: Linear Models

We used simple linear regression to examine the effects of daily heat index and daily maximum temperature on the start speed of four-seam fastballs while also considering the effects of the individual pitcher, inning, and ballpark location (home team)(AD1).



```{r filter_ff}


# Filter dataframe to only four-seam fastballs
pitch_tidy_ff <-
  pitch_tidy_db %>% 
  filter(pitch_type == "FF")

```



### Pitch speed and heat index  

```{r lm_heat}
pitch_tidy_ff %>% 
  lm(start_speed ~ heat_index, data = .) 
  broom::tidy() %>% 
  knitr::kable(digits = 3)
```


A one degree increase in heat index increased the start speed of four-seam fastballs by 0.036 mph (p < 0.001). This is a statistically significant difference, but it is not meaningful.  



### Pitch speed and maximum daily temperature

```{r lm_tmax}
pitch_tidy_ff %>% 
  lm(start_speed ~ tmax, data = .) %>% 
  broom::tidy() %>% 
  knitr::kable(digits = 3)
```

A one degree increase in maximum daily temperature increased the start speed of four-seam fastballs by 0.037 mph (p < 0.0001). This result is nearly identical to the heat index model.  


### Pitch speed by inning  


```{r lm_inning}

pitch_tidy_ff %>% 
  lm(start_speed ~ heat_index, data = .) %>%
  broom::tidy() %>% 
  knitr::kable(digits = 3)

```

.   

A one inning increase is associated with a 0.197 mph increase in the start speed of four-seam fastballs (p < 0.001). This may be due to closers coming into the game and throwing more heat. Two-tenths of a mph most likely does not make a difference on a per at-bat basis. 



 

### Pitch speed by inning, controlled for pitcher (top 200 fastball throwers)
```{r lm_inning_pitcher}

pitch_tidy_ff %>% 
  filter(pitcher_name != "NA") %>% 
  group_by(pitcher_name) %>% 
  summarise(n_ff = n()) %>% 
  arrange(desc(n_ff)) %>% 
  slice(1:200) %>%
  ungroup() %>% 
  left_join(., pitch_tidy_ff, by = "pitcher_name") %>% 
  lm(start_speed ~ inning + pitcher_name, data = .) %>%  
  broom::tidy() %>% 
  slice(1:5) %>% 
  knitr::kable(digits = 3) 
  
```

After adjusting for pitcher, the effect of inning is only an increase of 0.002 mph per inning (p = 0.03). The effect of inning will not be carried forward to further weather analysis.



### Pitch speed and heat index, controlled for home team 
```{r lm_speed_heat_home}

pitch_tidy_ff %>% 
  lm(start_speed ~ heat_index + home, data = .) %>%  
  broom::tidy() %>% 
  knitr::kable(digits = 3)

```

A one inning increase is associated with a 0.197 mph increase in the start speed of four-seam fastballs (p < 0.001). This may be due to closers coming into the game and throwing more heat. Two-tenths of a mph most likely does not make a difference on a per at-bat basis. 





### Future statistical analysis  

Additional analysis could examine non-linear effects of weather on pitch speed. Different approaches could include segmented regression or generalized additive models to test weather parameter thresholds.     


###Scatterplots

#### Speed and Heat index
```{r speed_heat_plot, warning = FALSE}
pitch_tidy_ff %>% 
  ggplot(., aes(x = heat_index, y = start_speed)) + 
    geom_point(aes(alpha = 0.5)) + 
    geom_smooth(method = "lm", color = "green") +
    geom_smooth(color = "black") +
    labs(
      title = "Heat index and fastball speed",
      x = "pitch speed (mph)",
      y = "heat index (deg C)"
    ) 


```


#### Speed and Inning
```{r inning_heat_plot, warning = FALSE}
pitch_tidy_ff %>% 
  ggplot(., aes(x = inning, y = start_speed)) + 
    geom_point() + 
    geom_smooth(method = "lm", color = "green") +
    geom_smooth() +
    labs(
        title = "Inning and fastball speed",
        x = "inning",
        y = "pitch speed (mph)"
      ) 
```


#### Speed and Heat index, by home stadium
```{r speed_heat_home_plot, warning = FALSE}
pitch_tidy_ff %>% 
  ggplot(., aes(x = heat_index, y = start_speed, color = home)) + 
    geom_point() + 
    geom_smooth(method = "lm") +
    labs(
          title = "Heat index and fastball speed, by home stadium",
          x = "heat index (deg C)",
          y = "pitch speed (mph)"
        ) 


```




